### 2022-06-27 tex/notes.tex
just pushing everything

\documentclass[twoside]{article}

\usepackage{aistats2021}
\usepackage{amsmath}
% If your paper is accepted, change the options for the package
% aistats2021 as follows:
%
%\usepackage[accepted]{aistats2021}
%
% This option will print headings for the title of your paper and
% headings for the authors names, plus a copyright note at the end of
% the first column of the first page.

% If you set papersize explicitly, activate the following three lines:
%\special{papersize = 8.5in, 11in}
%\setlength{\pdfpageheight}{11in}
%\setlength{\pdfpagewidth}{8.5in}

% If you use natbib package, activate the following three lines:
%\usepackage[round]{natbib}
%\renewcommand{\bibname}{References}
%\renewcommand{\bibsection}{\subsubsection*{\bibname}}

% If you use BibTeX in apalike style, activate the following line:
%\bibliographystyle{apalike}

\begin{document}

% If your paper is accepted and the title of your paper is very long,
% the style will print as headings an error message. Use the following
% command to supply a shorter title of your paper so that it can be
% used as headings.
%
%\runningtitle{I use this title instead because the last one was very long}

% If your paper is accepted and the number of authors is large, the
% style will print as headings an error message. Use the following
% command to supply a shorter version of the authors names so that
% they can be used as headings (for example, use only the surnames)
%
%\runningauthor{Surname 1, Surname 2, Surname 3, ...., Surname n}

% Supplementary material: To improve readability, you must use a single-column format for the supplementary material.
\onecolumn
\aistatstitle{Notes}

\section{Introduction}

\section{Constraints}
\subsection{Bounded Constraints}
\subsection{Simplex Constraints}
\subsection{Definite Matrix constraints}
\section{Softmax Transform}

The softmax function can be understood from Multinomial Logistic Regression employed for predicting probabilities of a Categorically distributed variable. Geometrically it maps $R^K$ to the boundary of a unit K-simplex(it is simply the convex hull of $k+1$ affinely independent points in $R^K$. Essentially it transforms a vector of size K to another vector of size K where the outputs sum to 1. It is worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so when a vector of size $K$ is transformed the $K_th$ vector is simply $1$- sumofk-1 vectors
R

\end{document}