### 2022-07-11 tex/paper.tex
add stickbreaking transform

\documentclass[11pt]{article}

\usepackage[letterpaper, total={6.5in, 9in}]{geometry}
\usepackage{cite}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathpazo}
\usepackage{sourcecodepro}

\newcommand{\setcomp}[2]{\left\{ #1 \ \Big|\ #2 \right\}}
\newcommand{\rngto}[1]{1{:}#1}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\absdet}[1]{\abs{#1}}

\begin{document}

\title{Efficient Unconstraining
  Parameter Transforms for Hamiltonian Monte Carlo}
\author{Meenal Jhajaria \\ \small Flatiron Institute \and Seth Axen \\
  \small University of T\"ubingen \and Bob
  Carpenter \\ \small Flatiron Institute}
\date{DRAFT: \today}
\maketitle

\begin{abstract}
  \noindent
  This paper evaluates the the statistical and computational
  efficiency of unconstraining parameter transforms for Hamiltonian
  Monte Carlo sampling.
\end{abstract}

\section{Introduction}

In statistical computing, we often need to compute high-dimensional
integrals over densities $\pi(x)$ (e.g., Bayesian estimation or
prediction, $p$-value calcuations, etc.).  The only black-box
techniques that work for general high-dimensional integrals are
Markov chain Monte Carlo (MCMC) methods.  The most effective MCMC
method in high dimensions is Hamiltonian Monte Carlo (HMC).  HMC works
by simulating the Hamiltonian dynamics of a fictitious particle
representing the value being sampled coupled with a momentum term.

Although it's possible to write HMC samplers that work for simply
constrained values such as upper- and/or lower-bounds
\cite{neal2011mcmc} or unit vectors \cite{byrne2013geodesic}, it is
much more challenging to do the same for complex constraints such as
simplexes or positive definite matrices or for densities involving
multiple constrained values.  Instead, it is far more common to map
the constrained values to unconstrained values before sampling
\cite{}.  This presents the issue of selecting which transform to use
among an infinite set of options, which is the topic of this paper.

Sampling algorithms play an important role in Statistics, in this case
we refer to drawing samples from a probability density or distribution
(not survey sampling). Monte Carlo Markov chain methods are commonly
used for high dimensional sampling. In this paper, we look at
Hamiltonian Monte Carlo, which uses hamiltonian dynamics to propose
samples for the Metropolis algorithm. Bayesian Inference often uses
parameters with constraints, but HMC optimizes on the unconstrained
space. It is difficult to use Monte Carlo estimation on a constrained
domain ~\cite{neal2008optimal}

\section{Changes of Variables}

If $X$ is a random variable with density $p_X$ and $Y = f(X)$ for a
smooth and bijective function $f$, then
\[
  p_Y(y) = p_X(f^{-1}(y)) \absdet{J_{f^{-1}}(y)},
\]
where the Jacobian of the inverse transform is defined by
\[
  J_{f^{-1}}(y) = \frac{\partial}{\partial y} \, f^{-1}(y)
\]
and
\[
  \absdet{J_{f^{-1}}(y)}
  = \abs{\det J_{f^{-1}}(y)}.
\]

\section{Transforms}

To draw samples from a distribution $p_X(x)$, defined on a constrained space $\mathcal{X}$ that can be uniquely parameterized by $n$ real degrees of freedom, we instead perform sampling in an unconstrained space $\mathcal{Y}=\mathbb{R}^m$ for $m \ge n$.

Let $f\colon \mathcal{Y} \to \mathcal{X}$ be a smooth and continuous map.
Only when $m = n$ can $f$ be bijective.
When $m > n$, multiple values of $y \in \mathcal{Y}$ may map to a given value of $x = f(y)$.
In this case, to make the function bijective, we consider an additional smooth and continuous map $g\colon \mathcal{Y} \to \mathcal{Z}$, where $\mathcal{Z}$ can be uniquely parameterized by $m - n$ real degrees of freedom.
We then define $h\colon \mathcal{Y} \to \mathcal{X} \times \mathcal{Z}: y \mapsto (f(y), g(y)) = (x, z)$.
$f$ and $g$ must be chosen so that $h$ is bijective almost-everywhere;
that is, the values of $y$ for which $h$ is non-bijective must have no probability mass.
When $h$ is non-bijective at single points, we call these singularities.
While singularities themselves are in general not a problem for sampling with MCMC, the region of a target density near the singularity tends to have high curvature.

Given a proper joint distribution with a smooth and continuous density $p_{X,Y}(x, y | \theta)$ for some parameters $\theta$, we can use the usual change of variables formula to write

\[
  p_Y(y | \theta) = p_{X,Z}(f(y), g(y) | \theta) |J_h(y)|.
\]

To sample from a target density $p_X(x | \theta)$, let $p_{X,Z}(x, z | \theta) = p_X(x | \theta) p_Z(z | x, \theta)$.
Then we must choose some proper distribution $p_Z$ that is smooth and continuous.
Given that choice, we find that
\[
  p_Y(y | \theta) = p_X(f(y) | \theta) p_Z(g(y) | f(y), \theta) |J_h(y)|.
\]

Note that when $m > n$, it is insufficient to specify a transform $h$.
One must also select a prior $p_Z$, and it is desirable to establish some general properties that a given pair $h$ and $p_Z$ should have.
In addition to the above properties, the optimal pair would
1. discard as few degrees of freedom $|m - n|$ as possible.
2. produce a $p_Z$ and $|J_y|$ and their gradients that are efficient to compute.
3. produce a $p_Y$ that is easy to sample from.

We can expand a bit on the latter point.
As a general rule, the more normal $p_Y$ is, the easier it is to sample from (REF?).
Additionally, near singularities, small perturbations to $y$ tend to result in larger changes to $h(y)$ than far from the singularity.
This warps the volume around the singularity, which causes $|J_y(y)|$ to smoothly approach infinity near the singularity.
If $m$ is large, this is generally not a problem, as the "curse of dimensionality" works in our favor, and there is exponentially more volume away from the singularity than near the singularity.
However, for low $m$, it is necessary that $p_{X,Z}$ goes to 0 near the singularity faster than $|J_y(y)|$ goes to infinity.
While $p_X$ is generally fixed by the user-specified model and may not have this necessary property, the user may then choose a $p_Z$ that has this property.

\section{Unit simplex}

A unit $N$-simplex is an $N + 1$-dimensional vector of non-negative
values that sums to one.  As such, there are only $N$ degrees of
freedom, because if $x$ is an $N$-simplex, then
\[
  x_N = -(x_1 + x_2 + \cdots + x_{N-1}).
\]
Simplexes are useful for representations of multinomial probabilities
(e.g., probabilities of categories in a classification problem).

The set of unit $N$-simplexes is conventionally denoted
\[
  \Delta^N = \setcomp{x \in \mathbb{R}^{N + 1}}{\textrm{sum}(x) = 1
    \textrm{ and }
    x_n \geq 0 \textrm{ for } n \in \rngto{N}}
\]
Geometrically, an $N$-simplex is the convex closure of $N+1$ points
that are 1 in one coordinate and 0 elsewhere.  For example, the
3-simplex is the complex closure of
$\begin{bmatrix}1 & 0 & 0 \end{bmatrix},
\begin{bmatrix} 0 & 1 & 0 \end{bmatrix}$,
and $\begin{bmatrix} 0 & 0 & 1 \end{bmatrix}$.

\subsection{Additive log ratio transform}

The unconstraining transform for the identified softmax is known as
the additive log ratio (ALR) transform \cite{aitchison???}, which is a
bijection $\textrm{alr}:\Delta^{N-1} \rightarrow \mathbb{R}^{N-1}$ defined by
\[
  \textrm{alr}(x)
  = \begin{bmatrix}\displaystyle
    \log \frac{x_1}{x_N} \cdots \log \frac{x_{N-1}}{x_N}
  \end{bmatrix}.
\]
The inverse additive log ratio transform maps values in $\mathbb{R}^{N-1}$ to
$\Delta^{N-1}$ by
\[
  \textrm{alr}^{-1}(x) = \textrm{softmax}(\begin{bmatrix}x &  0\end{bmatrix}),
\]
where for $u \in \mathbb{R}^N$,
\[
  \textrm{softmax}(u) = \frac{\exp(u)}{\textrm{sum}(\exp(u))}.
\]
To calculate the change of variables adjustment, we consider only the
first $N-1$ coordinates of the result, because the last is defined in
terms of the first $N-1$.  The function operating on the first $N-1$
coordinates is $s:\mathbb{R}^{N-1} \rightarrow \mathbb{R}^{N-1}$, defined by
\[
  s(x) = \frac{\exp(x)}{\exp(x) + 1}.
\]
The change of variables is then defined so that if $p_X(x)$ is a
density over simplices $x \in \Delta^{N-1}$, then we inverse transform
to an unconstrained density defined over $Y \in R^{N-1}$ by
\[
  p_Y(y) = p_X(\textrm{alr}^{-1}(y)) \absdet{J_{\textrm{alr}^{-1}}}.
\]

\subsection{StickBreaking Transform}

The StickBreaking transform ican be be understood from the stick-breaking construction for Dirichlet \cite{sethurman}. Intuitively, this comprises of recursively breaking a piece $x_i$ from a stick of unit length, where the leftover stick in the $i^{th}$ iteration is $ 1 - \sum_{1}^{i}x$. Let $y = f(x)$, then we define the stick-breaking mapping $ f \colon \Delta^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  R^{N-1}$, for $1 \leq i \leq N$ as:	
\[
y_i
= \mathrm{logit}(z_i) - \mbox{log}\left(\frac{1}{N-i}
   \right) \text{for break proportion} \, 
   z_i = \frac{x_i}{1 - \sum_{i' = 1}^{i-1} x_{i'}}.
\]

The inverse transform $ f^{-1} \colon R^{N-1} \xrightarrow{\makebox[0.4cm]{$\sim$}}  \Delta^{N-1}$ is defined as:
\[
x_i =
\left( 1 - \sum_{i'=1}^{i-1} x_{i'} \right) \text{for break proportion} \, z_i = \mathrm{logit}^{-1} \left( y_i
                             + \log \left( \frac{1}{N - i}
                                            \right)\right).
                                            \]

An $N$ dimensional unit simplex $\Delta^{N-1}$ has $N-1$ degrees of freedom (Notice that this inverse transform only maps the first $N-1$ elements, the last element of the simplex $x_{N} = 1 - \sum_1^{N-1}{x_i}$). The Jacobian matrix for $f^{-1}$ is a lower-triangular diagonal matrix. Much like the alr transform, for the change of variables we evaluate $\mathbf{J}_{i, i}$ where $i \in 1:N-1$.
\begin{align*}
\mathbf{J}_{i, i} &= \frac{\partial x_i}{\partial y_i}
=
\frac{\partial x_i}{\partial z_i} \,
\frac{\partial z_i}{\partial y_i}\\
\mathbf{J}_{i, i} &= \left(
  1 - \sum_{k' = 1}^{k-1} x_{k'}
   \right) z_k (1 - z_k),
\end{align*}

Absolute determinant of the diagonal matrix $\mathbf{J}$ is the product of its diagonal entries:
\begin{align*}
	\abs{\, det \, \textbf{J} \,} = \prod_{i=1}^{N-1} \textbf{J}_{i,i}
\end{align*}

The correction term $p_Y(y) = p_X(f^{-1}(y))\,
\prod_{i=1}^{N-1}z_i\,(1 - z_i)\left(1 - \sum_{i'=1}^{i-1} x_{i'}\right).$
\subsubsection{Softmax Transform}

The softmax function can be understood from Multinomial Logistic
Regression employed for predicting probabilities of a Categorically
distributed variable. Geometrically it maps $R^K$ to the boundary of a
unit K-simplex(it is simply the convex hull of $k+1$ affinely
independent points in $R^K$. Essentially it transforms a vector of
size K to another vector of size K where the outputs sum to 1. It is
worth noting that the mapping is actually from $R^K$ to $R^{K-1}$, so
when a vector of size $K$ is transformed the $K_th$ vector is simply
$1$- sumof k-1 vectors

\subsection{Simplex softmax parameterization}

Let $\Delta^n$ indicate the unit $n$-simplex with $n$ positive
elements that sum to 1 and $\Delta^n_-$ indicate the first $n-1$
elements, from which the final element can be uniquely determined.

We define the transformation
$\phi: \mathbb{R}^{n-1} \to \Delta^n_-: y \mapsto x_-$, where
$x=\begin{pmatrix}x_- \\ \frac{1}{r}\end{pmatrix} \in \Delta^n$,
$x_i = \frac{1}{r} e^{y_i}$ for $1 \le i \le n-1$, and
$r = 1 + \sum_{i=1}^{n-1} e^{y_i}$.

First we compute the scalar derivatives:
\[
\begin{aligned}
  \frac{\mathrm{d} r}{\mathrm{d} y_j}
  &= e^{y_j} = r x_j\\
  \frac{\mathrm{d} x_i}{\mathrm{d} y_j}
  &= \delta_{ij} \frac{1}{r} e^{y_i} - \frac{1}{r^2} e^{y_i}
  \frac{\mathrm{d} r}{\mathrm{d} y_j}
  = \delta_{ij} x_i - x_i x_j, \quad 1 \le i \le n-1
\end{aligned}
\]
where
$\delta_{ij} = \begin{cases} 1 &\text{if } i = j \\ 0
  &\text{otherwise}\end{cases}$ is the Kronecker delta.

If $\mathrm{diag}(x)$ is the diagonal matrix whose diagonal are the
elements of $x$, then the Jacobian is
\[
  J = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top) \operatorname{diag}(x_-),
\]
where $\boldsymbol{1}_n$ is the $n$-vector of ones.

Using Sylvester's determinant theorem,
$|I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = 1 -
\boldsymbol{1}_{n-1}^\top x_- = 1 - \sum_{i=1}^{n-1} x_i = x_n$, so
$$\mathrm{correction} = |J| = x_n \prod_{i=1}^{n-1} x_i = \prod_{i=1}^{n} x_i = \exp\left(\sum_{i=1}^{n-1} y_i\right) \left(1 + \sum_{i=1}^{n-1} e^{y_i}\right)^{-n}$$

\subsection{Simplex augmented softmax parameterization}

We define the transformation
$\phi: \mathbb{R}^n \to \Delta^{n-1} \times \mathbb{R}_{>0}: y \mapsto
(x_-, r)$, where $r = \sum_{i=1}^n e^{y_i}$ and
$x_i = \frac{1}{r} e^{y_i}$ for $1 \le i \le n-1$..

First we compute the scalar derivatives:
\[
\begin{aligned}
  \frac{\mathrm{d} r}{\mathrm{d} y_j}
  &= e^{y_j} = r x_j
  \\
  \frac{\mathrm{d} x_i}{\mathrm{d} y_j}
  &= \delta_{ij} \frac{1}{r} e^{y_i} - \frac{1}{r^2} e^{y_i} \frac{\mathrm{d} r}{\mathrm{d} y_j} = \delta_{ij} x_i - x_i x_j,
\end{aligned}
\]
which corresponds to the Jacobian
\[
  J = \begin{pmatrix}I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top & -x_- \\
    r \boldsymbol{1}_{n-1}^\top & r \end{pmatrix} \mathrm{diag}(x).
\]

For invertible $A$, the determinant of the block matrix
$\begin{pmatrix}A & B \\ C & D\end{pmatrix}$ is $|A| |D-CA^{-1}B|$.  A
square matrix is invertible iff its determinant is non-yero.  From the
previous section,
\[
  |I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top| = x_n > 0,
\]
so the determinant of the Jacobian is
\[
  |J| = x_n \left|r + r \boldsymbol{1}_{n-1}^\top (I_{n-1} - x_-
    \boldsymbol{1}_{n-1}^\top)^{-1} x_-\right|
  \prod_{i=1}^n x_i.
\]

Let $w = (I_{n-1} - x_- \boldsymbol{1}_{n-1}^\top)^{-1} x_-$. Then,
\[
\begin{aligned}
    w - x_- \sum_{i=1}^{n-1} w_i &= x_-\\
    w &= x_- \left(1 - \sum_{i=1}^{n-1} w_i\right)\\
    \sum_{i=1}^{n-1} w_i &= \sum_{i=1}^{n-1} \left( x_- (1 - \sum_{i=1}^{n-1} w_i) \right) = \left(\sum_{i=1}^{n-1} x_i \right) \left(1 - \sum_{i=1}^{n-1} w_i\right) = (1 - x_n)  \left(1 - \sum_{i=1}^{n-1} w_i\right)\\
    \sum_{i=1}^{n-1} w_i &= \frac{1 - x_n}{x_n} = \frac{1}{x_n} - 1\\
    w &= x_- \left(1 - \left(\frac{1}{x_n} - 1\right)\right) = \frac{1}{x_n} x_-
  \end{aligned}
\]

Then
\[
  |J| = x_n r \left|1 + \frac{1}{x_n}\sum_{i=1}^{n-1} x_i\right|
  \prod_{i=1}^n x_i = r \prod_{i=1}^n x_i
\]

To keep the target distribution proper, we must select a prior
distribution $\pi(r)$ for $r$.  If we choose $r \sim \chi_n$, then the
product of the correction and the density of the prior for $r$ is
proportional to

\[
  \mathrm{correction}
  = \pi(r) |J| = r^n e^{-r^2/2} \prod_{i=1}^n x_i
  = \exp\left(\sum_{i=1}^n y_i - \frac{1}{2}\left(\sum_{i=1}^n
      e^{y_i}\right)^2\right).
\]

Alternatively, if we choose $r \sim \mathrm{Gamma}(n, 1)$, then
\[
  \mathrm{correction} = \pi(r) |J| = r^n e^{-r} \prod_{i=1}^n x_i =
  \exp\left(\sum_{i=1}^n y_i - \sum_{i=1}^n e^{y_i}\right).
\]
This latter correction is equivalent to the sampling procedure from
the Dirichlet distribution with $\alpha_i=1$, where
$z_i \sim \mathrm{Exponential}(1)$ and
$y = \frac{z}{\sum_{i=1}^n z_i}$.

Both of these corrections can be captured with the generalization
\[
  \mathrm{correction}
  = \pi(r) |J|
  = r^n e^{-r^p/p} \prod_{i=1}^n x_i
  = \exp\left(\sum_{i=1}^n y_i - \frac{1}{p} \left(\sum_{i=1}^n e^{y_i}\right)^p\right),
\]
for $p > 0$, which corresponds to $r \sim \text{Generalized-Gamma}(1, n, p)$.

\subsubsection*{Acknowledgements}

We would like to thank $\ldots$.

\bibliography{all}{}
\bibliographystyle{plain}

\end{document}